WHOOMP! (There It Is) \\ Rapid Bayesian position reconstruction for gravitational\nobreakdashes-wave transients

### 2015-08-14 ms.tex
Define distance-independent amplitude

The distance integral is now performed differently from what we initially described in \cite{leo-singer-thesis,BerryLocalization}; the method described in the present work is about an order of magnitude faster. The innermost integral over distance $r$ may be written as
%
\begin{align}\label{eq:distance-integral}
    \mathscr{F} &=
        \int_{\mathrlap{r_\mathrm{min}}}^{\mathrlap{r_\mathrm{max}}}
        \exp\left[-\frac{p^2}{r^2}\right]
        I_0\left[\frac{b}{r}\right] r^m dr \nonumber \\
    &=
        \int_{\mathrlap{r_\mathrm{min}}}^{\mathrlap{r_\mathrm{max}}}
        \exp\left[-\frac{p^2}{r^2}+\frac{b}{r}\right]
        \overline{I}_0\left[\frac{b}{r}\right] r^m dr
\end{align}
%
or, completing the square,
%
\begin{align}
    \mathscr{F} &= \exp\left[\frac{p^2}{{r_0}^2}\right]
        \int_{\mathrlap{r_\mathrm{min}}}^{\mathrlap{r_\mathrm{max}}}
        \exp\left[-\left(\frac{p}{r} - \frac{p}{r_0}\right)^2\right]
        \overline{I}_0\left[\frac{2p^2}{r r_0}\right] r^m dr \\
    &= \exp\left[\frac{p^2}{{r_0}^2}\right] \mathscr{G},
\end{align}
%
where
%
\begin{align}
    p^2 &= \frac{1}{2} r^2 \sum_i {\rho_i}^2 \label{eq:likelihood-p-factor} \\
    b &= r \left| \sum_i \hat\rho_i \rho_i e^{i \tilde{\gamma}_i}
        a_i^*(\tilde{\tau}_i) \right| \label{eq:likelihood-b-factor} \\
    r_0 &= 2 p^2 / b \\
    \overline{I}_0(x) &= \exp(-|x|) I_0(x).
\end{align}
%
The coefficients $p^2$ and $b$ are nonnegative and independent of distance. $p$ has a maximum value of
%
\begin{equation}
    p_\mathrm{max} = \sqrt{\frac{1}{2} \sum_i \left(\frac{r_{1,i}}{r}\right)^2}.
\end{equation}
%
The symbol $\overline{I}_0$ denotes an exponentially scaled Bessel function. In the limit of large argument, $I_0(|x|) \sim \exp(|x|) / \sqrt{2 \pi |x|}$ \citep{Olver:2010:NHMF,NIST:DLMF}\footnote{\url{http://dlmf.nist.gov/10.40.E1}}. The scaled Bessel is useful for evaluation on a computer because it has a relatively small range $(0, 1]$ and varies slowly in proportion to $x^{1/2}$.

>>>>

The distance integral is now performed differently from what we initially described in \cite{leo-singer-thesis,BerryLocalization}; the method described in the present work is about an order of magnitude faster. We define $\rho_i = \omega_i / r$ in order to absorb all of the distance-independent terms in the amplitudes into $\omega_i$, and then define
%
\begin{align}
    p^2 &= \frac{1}{2} \sum_i {\omega_i}^2 \label{eq:likelihood-p-factor} \\
    b &= \left| \sum_i \hat\rho_i \omega_i e^{i \tilde{\gamma}_i}
        a_i^*(\tilde{\tau}_i) \right|. \label{eq:likelihood-b-factor} \end{align}
%
The innermost integral over distance $r$ may then be written as
%
\begin{align}\label{eq:distance-integral}
    \mathscr{F} &=
        \int_{\mathrlap{r_\mathrm{min}}}^{\mathrlap{r_\mathrm{max}}}
        \exp\left[-\frac{p^2}{r^2}\right]
        I_0\left[\frac{b}{r}\right] r^m dr \nonumber \\
    &=
        \int_{\mathrlap{r_\mathrm{min}}}^{\mathrlap{r_\mathrm{max}}}
        \exp\left[-\frac{p^2}{r^2}+\frac{b}{r}\right]
        \overline{I}_0\left[\frac{b}{r}\right] r^m dr
\end{align}
%
or, completing the square,
%
\begin{align}
    \mathscr{F} &= \exp\left[\frac{p^2}{{r_0}^2}\right]
        \int_{\mathrlap{r_\mathrm{min}}}^{\mathrlap{r_\mathrm{max}}}
        \exp\left[-\left(\frac{p}{r} - \frac{p}{r_0}\right)^2\right]
        \overline{I}_0\left[\frac{2p^2}{r r_0}\right] r^m dr \\
    &= \exp\left[\frac{p^2}{{r_0}^2}\right] \mathscr{G},
\end{align}
%
where
%
\begin{align}
    r_0 &= 2 p^2 / b \\
    \overline{I}_0(x) &= \exp(-|x|) I_0(x).
\end{align}
%
The coefficients $p^2$ and $b$ are nonnegative and independent of distance. $p$ has a maximum value of
%
\begin{equation}
    p_\mathrm{max} = \sqrt{\frac{1}{2} \sum_i \left(\frac{r_{1,i}}{r}\right)^2}.
\end{equation}
%
The symbol $\overline{I}_0$ denotes an exponentially scaled Bessel function. In the limit of large argument, $I_0(|x|) \sim \exp(|x|) / \sqrt{2 \pi |x|}$ \citep{Olver:2010:NHMF,NIST:DLMF}\footnote{\url{http://dlmf.nist.gov/10.40.E1}}. The scaled Bessel is useful for evaluation on a computer because it has a relatively small range $(0, 1]$ and varies slowly in proportion to $x^{1/2}$.
